---
title: "NYPD Shooting"
author: "D. Ikoma"
date: "2022-09-11"
output:
  pdf_document: default
  html_document: default
---

## Objective of analysis

New York is a quite attractive city for foreign travelers. But, gun shooting incidents are of critical concern for foreign travelers in New York. Therefore we would like to investigate the trend of gun shooting incidents in New York. Especially the trend of areas and time slots is helpful for travelers avoiding shooting incidents.

## Clear memory

At first, we clear memories in advance.

```{r clear_memory}
rm(list=ls())
gc();gc()
```

## Import libraries

We import libraries to use this analysis.

```{r import_libraries}
# ```{r import_libraries, echo = TRUE, message = FALSE}
Sys.setenv(LANGUAGE="en")
library(tidyverse)
library(lubridate)
```

## Read data

We read data from web site.

```{r get_data}
url_in <- "https://data.cityofnewyork.us/api/views/833y-fsy8/"
file_names <- c("rows.csv")
urls <- str_c(url_in, file_names)
data_NYPD <- read_csv(urls[1])
spec(data_NYPD)
```

## Tidy data

We investigate basic data and statistics. Next we remove some data and missing values for our analysis.

```{r tidy_data}
data_NYPD
summary(data_NYPD)
data_NYPD <- data_NYPD %>%
  mutate(OCCUR_DATE = mdy(OCCUR_DATE)) %>%
  select(-c(LOCATION_DESC, PERP_AGE_GROUP, PERP_SEX, PERP_RACE)) %>%
  drop_na(JURISDICTION_CODE)
data_NYPD
summary(data_NYPD)
```

We group shooting criminals counts by borough to analyze area trends.

```{r group_by_BORO}
data_by_BORO <- data_NYPD %>%
  group_by(BORO) %>%
  count(BORO) %>%
  summarize(cases = sum(n)) %>%
  ungroup()
data_by_BORO
```

Also, we group them by hour to know the dangerous time slots.

```{r group_by_hour}
tibble_opt <- list(
  "tibble.print_max" = 100,
  "tibble.print_min" = 20
)
options(tibble_opt)

data_by_hour <- data_NYPD %>% 
  mutate(hour = hour(OCCUR_TIME)) %>% 
  group_by(hour) %>%
  count(hour) %>%
  summarize(hours = sum(n)) %>%
  ungroup()
data_by_hour
```

## Visualize data

Here is a graphic representation of shooting counts by borough.

```{r visualize_by_BORO}
data_by_BORO %>%
  ggplot(aes(x = BORO, y = cases)) +
  geom_bar(stat = "identity") +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 0)) +
  labs(title = "NYPD shooting", y = "count")
```

We also visualize the shooting incidents counts by hour.

```{r visualize_by_hour}
data_by_hour %>%
  ggplot(aes(x = hour, y = hours)) +
  geom_bar(stat = "identity") +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 0)) +
  labs(title = "NYPD shooting", y = "count")
```

## Modeling

We model the number of criminal count by hour. We use a polynomial regression model.

```{r modeling}
mod <- lm(hours ~ hour + I(hour^2) + I(hour^3) + I(hour^4), data = data_by_hour)

summary(mod)
data_by_hour %>% slice_min(hours)
data_by_hour %>% slice_max(hours)
x_grid <- seq(0, 23)
new_df <- tibble(hours = x_grid)
data_by_hour_pred <- data_by_hour %>% mutate(pred = predict(mod))
data_by_hour_pred %>% ggplot() +
  geom_point(aes(x = hour, y = hours), color = "blue") +
  geom_line(aes(x = hour, y = pred), color = "red")

```

## Conclusions

Brooklyn has the highest number of gun crimes in New York City by borough. Brooklyn has the most population, so next we need to look at the number of crimes per capita. 


In addition, an analysis of the time period in which crimes occur shows that the number of crimes is low in the morning, with a minimum at 9:00 a.m., increasing in the evening and peaking at 11:00 p.m.


Regarding bias, it is necessary to investigate the effects of races, residents income, educational level, etc. with objective data and conduct an analysis that eliminates the bias.

```{r}
sessionInfo()
```